\LARGE \bf Efficient Sampling for Grasp Quality Evaluation on Gaussian Process Implicit Surface Representations [v-12 09-30-2014]

### 2014-09-30 paper.tex
changed everything

\title{\LARGE \bf
Efficient Sampling for Grasp Quality Evaluation with Gaussian Process Implicit Surface Representation }

\author{Michael Laskey, Zoe McCarthy, Florian T. Pokorny, Jeff Mahler, Sachin Patil,\\ Danica Kragic, Jur Van Den Berg, Pieter Abbeel, and Ken Goldberg}% <-this % stops a space

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

\todo{Write a better Intro, needs to have a better sell}
Many modern 3d sensors give noisy point clouds as output, so shape uncertainty is a common problem \cite{singhbigbird}. As shown in Fig. \ref{fig:noisy data}, the noise in our measurements of object shape can greatly change the surface normals and contact points, which are the parameters that most grasp metrics rely on.  A number of metrics have been proposed to evaluate form and force closure with scalar quality measures for grasping \cite{bicchi2000}.   However, only recently have people started looking into a metric's robustness to uncertainty.

We use Gaussian Processes \cite{rasmussen2006} to convert the point cloud measurements into an implicit surface with uncertainty in the shape. Gaussian Processes provide a formal way to incorporate uncertainty, however this sophisticated representation presents new challenges for evaluating the expected grasp quality. The wrench-space Ferrari-Canny force closure quality measure \cite{ferrari1992} calculates the maximum disturbance that can be resisted given bounds on the contact forces.

We assume that a robot's $m$ grippers approach the grasped object along straight lines (which we call lines of action), and analyze how this grasping model interacts with the probability distribution on shapes described by a Gaussian Process Implicit Surface (GPIS).
We calculate and visualize the theoretical probability distributions on grasp contact points and normals and the expected center of mass induced by the GPIS and line of action.
Then demonstrate how sampling from this explicit grasp distribution can be used to reduce the complexity associated with Monte-Carlo integration of the expected grasp quality over the distribution of shapes.

We also treat the problem of finding the best grasp in a set of proposed grasps as a n-arm bandit problem and propose an $\epsilon$ - greedy policy for finding the best grasp in the set in an anytime like manner.

\begin{figure}[ht!]
\centering
\includegraphics[scale = 0.3]{figures/Slide02.jpg}
\caption{An example of the noise from a Kinect-like sensor}
\vspace*{-10pt}
\label{fig:noisy data}
\end{figure}

\todo{Get an actual Kinect Image}

@@

While a variety of grasp metrics have been proposed, few explicitly looked at uncertainty in grasp parameters. Zheng et al. looked at friction and contact point uncertainty \cite{zheng2005}. He computed the maximum distance a contact point was allowed to deviate from its expected grasp, while maintaining force closure. Kehoe et al. looked at shape uncertainty for push grasps, however this was limited to parallel jaw grippers and only measured the probability of force closure \cite{kehoe2012toward}. More closely related to our approach is the work of Christopoulos and Schrater \cite{christopoulos2007handling}. They use a spline representation for uncertainty and measure the probability of losing force closure, their approach though is limited to two contacts. Our method looks at not only force closure, but provides the quality of the expected grasp and an upper bound on the expexted change from that quality.

The choice of using a Gaussian Process Implicit Surface to represent our uncertainty, stems from the fact that it provides a formal way to include various sources of noise in observations. Prior work has used uncertainty representation such as independent Gaussian noise on each vertex in a Polygonal mesh \cite{kehoe2012toward}. However, this doesn't provide a continuous function one can easily compute distributions on grasp parameters with. GPIS is also becoming more prevalent in the grasping community with recent work by Dragiev et al. \cite{dragiev2011} . The utilization of uncertainity around an implicit surface allows us to compute explicit distributions on contacts and surface normals, which to our knowledge has not been done formally.

\todo{Write about n-arm bandit literature}

\section{Gaussian Process Implicit Surfaces}

@@

\subsection{Gaussian Process (GP) Background}

Gaussian processes (GPs) are widely used in machine learning as a nonparametric regression method for estimating continuous functions from sparse and noisy data \cite{rasmussen2010gaussian}.
In a GP, a training set consists of input vectors $\mX = \{\bx_1, \ldots, \bx_n\}$, $\bx_i \in \mathbb{R}^d$, and corresponding observations $\by = \{y_1, \ldots, y_n\}$. The observations are assumed to be noisy measurements from the unknown target function $f$:
\begin{equation}
y_i = f(\bx_i) + \epsilon,
\end{equation}
where $\epsilon \sim \mN(0,\sigma^2)$ is Gaussian noise in the observations.
A zero-mean Gaussian process is completely specified by a covariance function $k(\cdot,\cdot)$, also referred to as a kernel.
Given the training data $\mD = \{\mX, \by\}$ and covariance function $k(\cdot,\cdot)$, the posterior density $p(f_*|\bx_*,\mD)$ at a test point $\bx_{*}$ is shown to be \cite{rasmussen2010gaussian}:
\begin{align}
	p(f_*|\bx_*,\mD) &\sim \mN\big(\mu(\bx_*), \Sigma(\bx_*)\big) \label{eq:GPposterior} \\
	\mu(\bx_*) &= k(\mX,\bx_*)^{\intercal}(K + \sigma^2I)^{-1}\by \label{eq:GPmean} \\
	\Sigma(\bx_*) &= k(\bx_*,\bx_*)-k(\mX,\bx_*)^{\intercal}(K+\sigma^2I)^{-1}k(\mX,\bx_*)\big) \label{eq:GPvar}
\end{align}
where $K \in \mathbb{R}^{n \times n}$ is a matrix with entries $K_{ij} = k(\bx_i,\bx_j)$ and $k(\mX,\bx_*) = [k(\bx_1,\bx_*),\ldots,k(\bx_n,\bx_*)]^{\intercal}$. 
This derivation can also be used to predict the mean and variance of the function gradient by extending the kernel matrices using the identities \cite{solak2003derivative}:

\begin{align*}
	\text{cov}\left(f(\bx_i), f(\bx_j) \right) &=  k(\bx_i, \bx_j) \\
	\text{cov}\left(\frac{\partial f (\bx_i)}{\partial x_k}, f(\bx_j) \right) &= \frac{\partial}{\partial x_k} k(\bx_i, \bx_j) \\
	\text{cov}\left(\frac{\partial f (\bx_i)}{\partial x_k}, \frac{\partial f (\bx_j)}{\partial x_l} \right) &= \frac{\partial^2}{\partial x_k \partial x_l} k(\bx_i, \bx_j)
\end{align*}

\subsection{Kernel Selection}
The choice of kernel is application-specific, since the function $k(\bx_i,\bx_j)$ is used as a measure of correlation between states $\bx_i$ and $\bx_j$. A common choice is the squared exponential kernel:
\begin{equation}
	k(\bx_i,\bx_j) = \nu^2\exp(-\frac{1}{2}(\bx_i - \bx_j)^{\intercal}\Lambda^{-1}(\bx_i - \bx_j))
\end{equation}
where $\Lambda= \text{diag}(\lambda_1^2,\ldots,\lambda_d^2)$ are the characteristic length scales of each dimension of $\bx$ and $\nu^2$ describes the variability of $f$.
Other common kernels relevant to GPIS are the thin-plate splines kernel \cite{williams2007} and the Matern kernel \cite{bjorkman2013enhancing}.

The measurement noise parameter $\sigma$ is usually estimated based on a model of the sensor used to collect measurements
The vector of remaining hyperparameters $\boldsymbol{\theta} = \{\nu, \lambda_1,\ldots,\lambda_d\}$ is optimized during the training process by maximizing the log-likelihood $p(\by|\mX,\boldsymbol{\theta})$, usually using gradient descent \cite{rasmussen2010gaussian}.
The log-likelihood function is subject to local maxima, and therefore the hyperparamter search often involves a search over the maxima found from several random initializations of the hyperparamters.

\begin{figure}[ht!]
\centering
\includegraphics[scale = 0.3]{figures/Slide03.jpg}
\caption{Left: A surface represented as a Truncated Signed Distance Field (TSDF).
Right: A GPIS reconstruction from noisy samples of the left surface's TSDF.}
\vspace*{-10pt}
\label{fig:GPIS_TSDF}
\end{figure}
%\section{Related Work}

\section{Problem Definition}

%The objective of our algorithm is to provide a quality of the expected grasp, $Q_l^-(\bar{g})$ \cite{ferrari1992}, and a lower bound $b$ that is the expected deviation from the quality of the expected grasp.

We assume a bounded 2-d rectangular workspace $\mathcal{R}$.
We parameterize a single grasp $g$ on an object with the following tuple $g = ( \textbf{c}_1,...,\textbf{c}_m,\textbf{n}_1,...,\textbf{n}_m,\textbf{z},\tau )$.
We have an indexing set $I$ of $m$ point contacts and surface normal on the object: for $i \in I$ the contact is located at $c_i$ with surface normal $n_i$.
The object has a center of mass $z$ and friction coefficient $\tau$.
The line segment $\gamma(\cdot)$ has endpoints $a,b$ that are defined as the start of the gripper and the intersection of the line with the end of the workspace respectively, as shown in Fig. 
 \ref{fig:line_of_action}.

We assume that the robot grippers move along the line of action until they make contact with the object and stop.
Since we have uncertainty in the shape this induces a distribution on the grasp parameters subject to the line, hence $p(g) = ( p(\textbf{c}_1),...,p(\textbf{c}_m),p(\textbf{n}_1),...,p(\textbf{n}_m),\bar{\textbf{z}},\tau )$.
We note that $\tau$ is considered known and $\bar{z}$ is the expected center of mass, not a full probability distribution. In section \ref{sec:distgrasp}, we demonstrate how to efficiently compute the distributions on contact points and surface normals and expected center of mass.

@@

 For the following derivations we introduce the following,
 $\theta(x) = ( \mu(x),\Sigma(x) )$, where $\theta(x)$ is a tuple consisting of the mean and covariance functions given by the trained GPIS model \cite{rasmussen2006} .

 To calculate $p(g)$, we assume a gripper contacts approaches along a parameterized line of action, or a 1-dimensional curve in the work space, defined by $\gamma(t)$. See Fig \ref{fig:line_of_action}, for a detailed illustration. Each gripper contact is defined by a line of action, so we assume the following tuple is provided $\Gamma = ( \gamma_1(\cdot),...,\gamma_m(\cdot) )$, these approach trajectories are then used to compute a distribution on grasp parameters.

 %We also define the  level set function as $f(x): \mathbb{R}^d \rightarrow \mathbb{R}$ and an implicit surface as $\mathcal{S} = \{ x \ | \ f(x) = 0 \}$.

\todo{Get a better figure for line of action model}

\begin{figure}[ht!]
\centering
\includegraphics[scale = 0.3]{figures/Slide01.jpg}
\caption{Parameterized Line of Action along an object}
\vspace*{-10pt}
\label{fig:line_of_action}
\end{figure}
%\textbf{TODO:MORE DESCRIPTIVE GRAPHIC}
\subsection{Distribution on Contact Points} 
In our implementation we discretize along the line $\gamma(t)$ evenly but write the derivation in continuous form for generality.
The probability distribution along the line $\gamma(t)$ is given by the following:

\begin{equation}
p\big(f(\gamma(t))|\theta(\gamma(t)): \forall t \in [a,b] \big) 
=
\mN(\mu_{a:b},\Sigma_{a:b}).
\end{equation}

This gives the signed distance function distributions along the entire line of action in the workspace as a multivariate gaussian.
We would like to find the distribution on the first contact point, which we can define as when the signed distance function $f(\gamma(t))$ is $0$ and all previous times $\tau$ we have $f(\gamma(\tau)) > 0$ for $0 \leq \tau < t$.
This ensures $t$ is at the edge of the surface (having $f(\gamma(t)) = 0$) and that for all previous $\tau$ the gripper was outside of the surface (having $f(\gamma(\tau)) > 0$).
We thus compute this as the joint distribution $p\big(\textbf{c}_i= \gamma(t)\big) = p\big(f(\gamma(t))=0, f(\gamma(\tau))> 0: \forall \tau \in [0,t)\big)$.
%Hence we want to know at a point what the distribution  is that it is at the surface and the probability that all points before are before the surface,
This avoids the problem of the distribution producing multiple modes along the line: one for each intersection with the surface.
We now derive this distribution

\[
  p\big(\textbf{c}_i = \gamma(t)\big) \propto 
\]
\[
  p\big(f(\gamma(t)) = 0\big)P\big(f(\gamma(\tau)) > 0 | f(\gamma(t)) = 0: \forall \tau \in [0,t)\big)
\]

where we only indicate proportionality and will later normalize to probability $1$.
Using the first product in the equation can be computed easily using the marginalization of a multivariate Gaussian distribution and the second one can be rewritten by conditioning the distribution \cite{petersen2008matrix}.

\begin{align*}
p_c\big(f(\gamma(\tau)): \forall \tau \in [0,t)\big) = p\big(f(\gamma(\tau))  | f(\gamma(t)) = 0\big)  
\end{align*}

The following can now be said:

\[
  p\big(\textbf{c}_i = \gamma(t)\big) = 
\]
\[
  = \frac{1}{\eta} p(f(\gamma(t)) = 0) P_c\big(f(\gamma(\tau)) > 0): \forall \tau \in [0,t)\big)			 
\]

for the appropriate $\eta$ normalization factor.
The second product term can be evaluated by calculating the cumulative distribution of a multivariate Gaussian, which we calculate with the Matlab function $mvncdf$.
We show again the theoretical distribution on $\textbf{c}_i$ calculated for a given GPIS and approach direction in Fig.
\ref{fig:GraspContactPt}.
%efficiently evaluated by looking up the cumulative distribution of a multivariate Gaussian, which is common in most software packages \cite{matlab}.

>>>>

\title{\LARGE \bf
Efficient Sampling for Grasp Quality Evaluation on Gaussian Process Implicit Surface Representations [v-12 09-30-2014]}

\author{Michael Laskey, Zoe McCarthy, Jeff Mahler, Florian T. Pokorny, Sachin Patil,\\  Danica Kragic, Pieter Abbeel, and Ken Goldberg}% <-this % stops a space

@@

\begin{abstract}
Gaussian Process Implicit Surface (GPIS) have emerged recently as a new Bayesian shape representation.  Former methods for grasp evaluation with shape uncertainty require sampling from  distributions on shape. The standard approach of drawing samples from a GPIS in 2D by means of a evaluation on a uniform grid has a complexity of $O(n^6)$(2D) and $O(n^9)$ in the number of grid steps $n$.. We propose an alternative approximation to sampling from shape distributions and demonstrate how that reduce complexity from $O(n^6)$ in 2D and $O(n^9)$ in 3D to $O(n^3)$ for evaluation of a single grasp. We furthermore introduce the idea of adaptive sampling in this context by formulating it as a multi-arm bandit problem for speeding up the selection of the highest quality grasp in a set of proposed grasps. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

\todo{Get High res GPIS visualizations, Incorporate next round of feedback}
Consider a robot packing boxes in a shipping warehouse environment. The robot will frequently encounter objects that it has not seen before. The robot will have to plan stable grasps without any prior information other than its sensor data. As seen in Fig. \ref{fig:noisy data}, with today's RGB and depth sensors, surface properties, such as specularity and transparency, can greatly effect the returned observations. Grasp planners based on deterministic algorithms are not designed to handle such noise and could fail. To plan grasps in a partially observed world, recent research has focused on incorporating uncertainity in the shape of the object \cite{kehoe2012estimating} \cite{kehoe2012estimating} \cite{christopoulos2007handling} \cite{zheng2005}..

While past work has encoded uncertainty in a polygonal mesh representation \cite{kehoe2012estimating}, an alternative is to work with the signed distance fields (SDFs), a continuous valued function that is an implicit surface: zero-valued at the surface of the object, positive value outside the surface and negative valued inside the surface. This representation is growing in popularity in the robotics field for its ability to integrate range data in real time \cite{newcombe2011kinectfusion}\cite{curless1996volumetric}. To incorporate uncertainty one can use a Gaussian Process Implicit Surface (GPIS), a Gaussian process over signed distance field. Not only does GPIS present a straightforward and formal way to incorporate measurements from visual, laser and tactile sensors, they  also allow for a continuous differentiable function on shape, which can be used for a feedback grasp controller, finding locally optimal grasps and active exploration  \cite{dragiev2011}\cite{jeffs}\cite{hollinger2013}.

\begin{figure}[ht!]
\centering
\includegraphics[width=8.5cm,height=4cm]{figures/Slide02.jpg}
\caption{\footnotesize
Left: HD image of a common measuring cup from in a household environment. Right: Image taken from PR2's Primesense Carmine sensor, large parts of the measuring cup are not detected by the depth sensor and are not visible in the RGB image.}
\vspace*{-10pt}
\label{fig:noisy data}
\end{figure}

Given a method to represent shape uncertainty, a common technique to evaluate grasp quality is to sample from the distribution on shape and perform Monte-Carlo integration on a given grasp metric \cite{kehoe2012estimating}\cite{kehoe2012toward}\cite{christopoulos2007handling}. Applying this technique to GPIS requires an exhaustive sampling of the signed distance field, which scales $O(n^6)$ for a 2D $n \times n$ workspace and $O(n^9)$ for a 3D workspace.The computationally complexity of sampling is derived from the fact sampling from a Gaussian requires matrix inversion of the entire signed distance field. Our paper reduces this complexity to $O(n^3)$ by using the insight that for grasp quality evaluation one only cares about the parameters that define a grasp: contacts points, surface normals, center of mass and friction coefficient \cite{pokorny2013classical}. We constrain the contact points by considering the approach trajectory of a gripper.  Hence, sampling the whole signed distance field is not needed, but instead one can only work with the distributions on grasp parameters given the shape uncertainty model and gripper trajectory. In light of this our contributions are the following:

\begin{enumerate}
	\item Derivation of  distributions on grasp parameters given GPIS shape uncertainty model and approach trajectory of gripper.
	\item  Demonstrating both theoretically and empirically how those distributions reduce the computationally complexity of sampling.
	\item Applying adaptive sampling techniques to efficiently find high quality grasps in a given set of proposed grasps. 
\end{enumerate}

@@

 Past work on grasping under uncertainty has focused on state uncertainty \cite{goldberg1990bayesian}\cite{goldberg1990stochastic} \cite{stulp2011learning},  uncertainty in object pose \cite{christopoulos2007handling}  \cite{felip2009robust} \cite{kim2012physically} or uncertainty in the location of contact with the object \cite{zheng2005}. Robust grasp selection for parallel jaw grippers on uncertain shapes was studied by Kehoe et al. for both zero-slip push grasps \cite{kehoe2012toward} and slip push grasps \cite{kehoe2012estimating} using Monte Carlo integration for the expected grasp quality, but this approach was limited to polygonal shapes.  Christopoulo et al. sampled spline fits for 2-dimensional planar objects to measure the quality of uncertainty, however it is has not been shown how a spline based method could incorporate non-visual sensing information for representing shape uncertainty.

 The choice of using Gaussian Process Implicit Surfaces to represent uncertainty, is based on the fact it provides a formal way to include various sources of noise from observations such as tactile, laser and visual. Futhermore, implicit surface representations are becoming increasingly popular in robotics due to their success in real-time surface modeling from range images \cite{curless1996volumetric}\cite{newcombe2011kinectfusion}\cite{hornung2013octomap}. Some exciting work done with GPIS representation are the following: Hollinger et al. used GPIS to perform active sensing on the hulls in underwater boats \cite{hollinger2013}. Dragiev et al. showed how GPIS can represent shapes for grasps and used  as a grasp controller on the continuous signed distance function \cite{dragiev2011}. Mahler et al. used GPIS representation to find locally optimal grasps by framing grasp planning as an optimization problem \cite{jeff}.

  For grasp quality evaluation with shape uncertainty a common method involves sampling from the distribution on shape \cite{kehoe2012estimating} \cite{kehoe2012toward} \cite{christopoulos2007handling}. With GPIS this requires computationally intensive sampling from a signed distance field that is the dimension of the workspace. We propose an alternative distribution to sample from, based on the idea that we only need to consider the approach trajectory of the gripper.

Monte-Carlo integration involves drawing random samples from a distribution to approximate an integral \cite{caflisch1998monte}. While previous work has looked at quality evaluation of a single grasp with uncertainty \cite{christopoulos2007handling}\cite{zheng2005}, few have compared many different grasps with uncertainty and efficiently chosen the best one.  Adaptive Sampling is the idea of allocating samples based on previous observations or rewards, which can be described as a multi-arm bandit problem \cite{barto1998reinforcement}. Prior work by Kehoe et al. \cite{kehoe2012toward} used an adaptive sampling approach that removed low quality grasps based on values received, but this method had no statistical guarantees and could lead to high quality grasps being removed. To achieve guarantees, we frame the problem as a "budgeted multi-arm bandit" problem, where exploration and exploitation are decoupled and the agent has to allocate resources to explore for a finite time then decide which arm to exploit  \cite{madani2004budgeted}. Drawing from the literature on Monte-Carlo sampling we estimate the confidence bounds of grasps and prune only those that are statistically distinguishable \cite{caflisch1998monte}.

\section{Preliminaries and Problem Statement}

@@

\subsection{Gaussian Process (GP) Background}
GPs provide an infinite dimensional analogue of Multivariate
Gaussian Distributions which allow us to model distributions over
functions. Gaussian processes (GPs) are widely used in machine learning as a nonparametric regression method for estimating continuous functions from sparse and noisy data \cite{rasmussen2010gaussian}.
In a GP, a training set consists of input vectors $\mX = \{\bx_1, \ldots, \bx_n\}$, $\bx_i \in \mathbb{R}^d$, and corresponding observations $\by = \{y_1, \ldots, y_n\}$. The observations are assumed to be noisy measurements from the unknown target function $f$:
\begin{equation}
y_i = f(\bx_i) + \epsilon,
\end{equation}
where $\epsilon \sim \mN(0,\sigma^2)$ is Gaussian noise in the observations.
A zero-mean Gaussian process is completely specified by a covariance function $k(\cdot,\cdot)$, also referred to as a kernel.
Given the training data $\mD = \{\mX, \by\}$ and covariance function $k(\cdot,\cdot)$, the posterior density $p(f_*|\bx_*,\mD)$ at a test point $\bx_{*}$ is shown to be \cite{rasmussen2010gaussian}:
\begin{align*}
	p(f_*|\bx_*,\mD) &\sim \mN\big(\mu(\bx_*), \Sigma(\bx_*)\big) \\
	\mu(\bx_*) &= k(\mX,\bx_*)^{\intercal}(K + \sigma^2I)^{-1}\by \\
	\Sigma(\bx_*) &= k(\bx_*,\bx_*)-k(\mX,\bx_*)^{\intercal}(K+\sigma^2I)^{-1}k(\mX,\bx_*)\big) 
\end{align*}
where $K \in \mathbb{R}^{n \times n}$ is a matrix with entries $K_{ij} = k(\bx_i,\bx_j)$ and $k(\mX,\bx_*) = [k(\bx_1,\bx_*),\ldots,k(\bx_n,\bx_*)]^{\intercal}$. 
This derivation can also be used to predict the mean and variance of the function gradient by extending the kernel matrices using the identities \cite{solak2003derivative}:

\vspace{-2ex}
\begin{align}
	\text{cov}\left(f(\bx_i), f(\bx_j) \right) &=  k(\bx_i, \bx_j) \\
	\text{cov}\left(\frac{\partial f (\bx_i)}{\partial x_k}, f(\bx_j) \right) &= \frac{\partial}{\partial x_k} k(\bx_i, \bx_j) \label{eq:mean_gradient}\\
	\text{cov}\left(\frac{\partial f (\bx_i)}{\partial x_k}, \frac{\partial f (\bx_j)}{\partial x_l} \right) &= \frac{\partial^2}{\partial x_k \partial x_l} k(\bx_i, \bx_j)\label{eq:cov_gradient}
\end{align}

\begin{figure}[ht!]
\centering
\includegraphics[width = 8.5cm, height = 14cm ]{figures/Slide03.jpg}
\caption{\footnotesize Six example objects illustrating transparency, specularity, sensor noise that illustrate the need for shape uncertainty: (from top to bottom) tape, squirt bottle, stapler, loofa, marker, water bottle. Displayed from right to left are the HD color image, a point cloud observation from a Primesense Carmine, GPIS visualization technique from \cite{jeffs}, the nominal shape on a $25 \times 25$ grid.  }
\vspace*{-10pt}
\label{fig:GPIS_TSDF}
\end{figure}
%\section{Related Work}

For our kernel choice we decided to use the square exponential kernel, a major reason for this was its differentiability. Other kernels relevant to GPIS are the thin-plate splines kernel and the Matern kernel \cite{williams2007}.

We construct a GPIS by learning a Gaussian Process to fit measurements of a signed distance field of an unknown object.  Precisely, $x_i \in \mathbb{R}^2$ in 2D and $x_i \in \mathbb{R}^3$ in 3D, and $y_i \in \mathbb{R}$ is a noisy signed distance measurement to the unknown object at $x_i$.

\subsection{Line of Action}
Similar to the work of \cite{christopoulos2007handling}, we assume that each gripper finger approaches along a \textit{line of action}, a 1D curve $\gamma(t)$ with endpoints $a$ and $b$ as seen in Fig. \ref{fig:line_of_action}. A gripper finger starts at point $a$ and moves towards $b$, we assume $a$ is far enough away to be collision free of the object.  Each gripper contact is defined by a line of action, so we assume the following tuple is provided $\Gamma = ( \gamma_1(\cdot),...,\gamma_m(\cdot) )$, which designates a proposed \textit{grasp plan}.

\begin{figure}[ht!]
\centering
\includegraphics[scale = 0.3]{figures/Slide01.jpg}
\caption{Illustration of a grasp plan $\Gamma$ composed of two lines of action, $\gamma_1(t)$ and $\gamma_2(t)$}
\vspace*{-10pt}
\label{fig:line_of_action}
\end{figure}

\subsection{Problem Definition}

Given a 2-D workspace $\mathcal{W}$, with an unknown object represented as a trained GPIS model and set of possible grasp plans $G$. We are interested in determining

\begin{align}\label{eq:problem_def}
\Gamma^* \in \mbox{argmax}_{\Gamma \in G} E(Q(\Gamma))
\end{align}

with respect to a chosen grasp metric Q.

\section{Evaluating a Grasp}
In order to solve our problem definition, we must evaluate $E(Q(\Gamma))$ for a given grasp plan $\Gamma$. We will first discuss which grasp metric $Q$ we chose and then proceed into evaluating the expectation $E$ efficiently.

\subsection{Grasp Metric}
In their pioneering work over two decades ago Ferrari and Canny \cite{ferrari1992}, demonstrated a method to rank grasps by considering their contact points and surface normals. Importantly the magnitude of Q yields a measurement that allows one to rank grasps by their physical stability and evaluate the property of force-closure. Furthermore, it has wide spread use in grasp packages like GraspIT\cite{miller2004graspit}, OpenGrasp\cite{73} and Simox \cite{vahrenkamp2010simo}, which motivates studying its effect with uncertainties.

The $L^1$ version of the metric works by taking as input the contact points $\textbf{c}_1,...,\textbf{c}_m$, surface normals $\textbf{n}_1,...,\textbf{n}_m$, center of mass $z$ and friction coefficient $\mu$. Then constructing a convex hull around the wrenches made up of those parameters and finding the radius of the largest unit ball centered at the origin in wrench space. A wrench is defined as concatenation of a force and torque vector.  If the convex hull doesn't enclose the origin, the grasp is not in force-closure. Thus a grasp can be parameterized by the following tuple $g = \lbrace \textbf{c}_1,...,\textbf{c}_m,\textbf{n}_1,...,\textbf{n}_m,\mu, \textbf{z} \rbrace$, our method is applicable to all grasp metrics that represent a grasp as the tuple $g$, such as \cite{christopoulos2007handling}\cite{li1988task}.

\subsection{Calculating the Expected Grasp Quality}
Given a proposed grasp plan $\Gamma$, the expected grasp quality can be evaluated as follows:

\vspace{-2ex}
\begin{align}\label{eq:shape_sampling}
E(Q(\Gamma)) = \int Q(g|S,\Gamma) p(S) dS
\end{align}

Where $Q(g|S,\Gamma)$ is the grasp quality that is computed on a shape sample drawn from $p(S)$. To compute this we intersect the zero crossing of the level set with the propose grasp plan $\Gamma$ and determine the parameters $g$, this has been the approach taken in previous work  \cite{kehoe2012estimating} \cite{kehoe2012toward} \cite{christopoulos2007handling}. See Fig. \ref{fig:shape_samples} for an example of what samples drawn from $p(S)$ induced by GPIS look like. For computational reasons we approximate the integral via Monte-Carlo Integration. We use importance sampling to draw from the distribution induced by GPIS and calculate the following

\begin{figure}[ht!]
\centering
\includegraphics[width = 8.5cm, height= 5cm ]{figures/Slide13.jpg}
\caption{Shape samples drawn from Eq. \ref{eq:joint_shape} on the object in the upper left corner. Given a shape sample we highlight the zero-crossing of the level set in black}
\vspace*{-10pt}
\label{fig:shape_samples}
\end{figure}

\begin{align*}
E(Q(\Gamma)) \approx \sum_{i=1}^N Q(g|S_i,\Gamma) , S_i \in p(S)
\end{align*}

To compute the above distribution we must draw samples from $p(S)$. In order to draw shape samples from a GPIS, one needs to sample from signed distance function, $sd$, over the joint on all points in the workspace $\mathcal{W}$ or $p(sd(\mathcal{W}))$. Since this is a GPIS, we know the following

\begin{align}\label{eq:joint_shape}
p(S) = p(sd(\mathcal{W})) \sim N(\mu(\mathcal{W}),\Sigma(\mathcal{W}))
\end{align}

Thus if the workspace is an $n \times n$ grid, the joint distribution is an  $n^2$ multi-variate Gaussian, due to $sd:\mathbb{R}^2 \rightarrow \mathbb{R}$.  Sampling from a Gaussian involves inverting the covariance matrix and inversion is in the naive way $O(n^3)$ \cite{petersen2008matrix}. Thus the complexity of this operation is $O(n^6)$ in 2D and $O(n^9)$ in 3D.

To reduce complexity we propose sampling not from the shape distributions, but instead from the distributions on the grasps parameters themselves. We recall that a grasp according to our metric is defined as the tuple $g = \lbrace \textbf{c}_1,...,\textbf{c}_m,\textbf{n}_1,...,\textbf{n}_m,\mu, z \rbrace$. We are thus interested in calculating $p(g|\Gamma,\mu(x),\Sigma(x))$. The distribution on a grasp is defined then as:

\begin{align}\label{eq:joint_on_shape}
p(g) = p\big(\textbf{c}_1,...,\textbf{c}_m,\textbf{n}_1,...,\textbf{n}_m|\Gamma,\mu(x),\Sigma(x)\big)
\end{align}

We note here that we currently use the friction coefficient $\mu$ and the expected center of mass $\bar{z}$ as deterministic values. For grippers that do not approach along the same line of action (i.e. non-parallel jaw grippers) we make the assumption that each contact and normal pair is independent, or

\vspace{-2ex}
\begin{align}\label{eq:independence}
p(g) = &\prod_{i=1}^mp\big(\textbf{c}_i,\textbf{n}_i|\gamma_i(t),\mu(x),\Sigma(x)\big)
\end{align}

We compute the expected grasp quality now as follows:

\vspace{-2ex}
\begin{align}\label{eq:grasp_sampling}
E(Q(\Gamma)) = \sum_{i=1}^N Q(g_i) , g_i \in p(g)
\end{align}

We will now show how these distributions can be computed and how the complexity for evaluating a grasp is reduced from $O(n^6)$ to $O(n^3)$.

@@

 To sample from $p(g)$, we need to compute the distributions associated with a line of action $p(n_i|c_i)p(c_i|\gamma_i(t),\mu(x),\Sigma(x))$ on each of the parameters: contact points $c_i$, surface normals $n_i$  and expected center of mass $\bar{z}$. 
 .
\subsection{Distribution on Contact Points} 
The probability distribution along the line $\gamma(t)$ is given by:

\vspace{-2ex}
\begin{equation} \label{eq:line_of_act_dist}
p\big(sd(\gamma(t)) ; \mu(t),\Sigma(t)\big) \ \forall t \in [a,b] 
\end{equation}

This gives the signed distance function distributions along the entire line of action in the workspace as a multivariate Gaussian. One could think of this as a marginalization of all other points in signed distance field except the line of action. We would like to find the distribution on the first contact point, which we can define as when the signed distance function $sd(\gamma(t))$ is $0$ and all previous times $\tau$ we have $sd(\gamma(\tau)) > 0$ for all $\tau$ such that $0 \leq \tau < t$.We thus compute this as the joint distribution $p\big( \textbf{c}_i= \gamma(t)\big) = p\big(sd(\gamma(t))=0, sd(\gamma(\tau))> 0 \big)$ $  \forall \tau \in [0,t)$.

We now derive this distribution:

\vspace{-2ex}
\begin{align}
 & p\big(\textbf{c}_i = \gamma(t)\big) = \label{eq:contact_theory} \\
  & \frac{1}{\eta} p\big(sd(\gamma(t)) = 0\big)P\big(sd(\gamma(\tau)) > 0 | sd(\gamma(t)) = 0\big) \\
  & \forall \tau \in [0,t)
\end{align}

Where $\eta$ is a proportionality constant. The first product in the equation can be computed easily using the marginalization of a multivariate Gaussian distribution and the second term can be solved by first conditioning on the joint $p(sd(\gamma(t))|sd(\gamma(t)) = 0)$ and then computing the cumulative distribution function  \cite{petersen2008matrix}.  We show the theoretical distribution on $\textbf{c}_i$ calculated for a given GPIS and line of action in Fig.
\ref{fig:GraspDist}.

\begin{figure*}[ht!]
\centering
\includegraphics[width = 17cm, height = 4.5cm]{figures/Slide04.jpg}
\caption{ \footnotesize (Left to Right): Line of action for a given gripper on an uncertain surface representing a measuring cup. Distribution $p(c)$ as a function of $t$, the position along the line of action $\gamma(t)$. The two modes correspond to the different potential contact points, either the handle or the base of the cup. Lastly, the distribution on the surface normals (inward pointing) along $\gamma(t)$ described by equation \ref{eq:normal_dist}. }
\vspace*{-10pt}
\label{fig:GraspDist}
\end{figure*}

In practice Eq. \ref{eq:contact_theory} is not efficient to compute because it requires computation of the cumulative distribution function of a large multivariate Gaussian, so instead we sample from Eq. \ref{eq:line_of_act_dist} and find the first positive to negative zero-crossing of the sample along the line.

\subsection{Distribution on Surface Normals} 
Using Eq. \ref{eq:mean_gradient} and Eq. \ref{eq:cov_gradient}, we can compute the mean of the gradient $ \mu_{\nabla}(x)$ and the covariance of the gradient $\Sigma_{\nabla}(x)$ respectively. Thus we can compute the distribution around the surface normal for a given point in $\mathcal{W}$. We can now write

\vspace{-2ex}
\begin{align*}
p\big(\textbf{n}_i|\textbf{c}_i = \gamma(t)\big) = p\big(\textbf{n}_i |\mu(\gamma(t)), \Sigma(\gamma(t)) \big)
\end{align*}

One interesting effect of this technique is that we can now marginalize out the line of action model and visual what the surface normal distribution is along a given line of action. To our knowledge this is the first attempt to visual surface normals along a grasp plan. Marginalization can be performed as follows:

\vspace{-2ex}
\begin{equation}
p(\textbf{n}_i ) = \int_a^b   p\big(\textbf{n}_i = \textbf{v} | \textbf{c}_i = \gamma(t) \big)p\big(\textbf{c}_i = \gamma(t)\big) dt
\end{equation}
